# (PART\*) Part V: Conclusion {-}

# Summary

Implementing Bayesian methods has never been easier

<span class="pack">rstanarm</span> and <span class="pack">brms</span> make allow one to worry about the concepts instead of syntax

Bayesian analysis provides more ways to combat complex models, diagnosis issues, get creative with results

Even standard/traditional/basic models become more fun!




# Exercise

Whether a survey respondent agrees or disagrees with a conservative statement about the role of women in society

- 1974/1975 Survey
- Women should take care of running their homes and leave running the country up to men
- Modeled as a function of the gender and education of the respondents
- Data from the <span class="pack">HSAUR3</span> package



```{r basic_logistic_data_prep, echo=FALSE, eval=FALSE}
data("womensrole", package = "HSAUR3")
womensrole = womensrole %>% 
  mutate(total = agree + disagree,
         agree50 = agree/total >= .5)
readr::write_csv(womensrole, 'data/womensrole.csv')
```

```{r basic_logistic, echo=-2, eval=-c(2,4)}
data("womensrole", package = "HSAUR3")
womensrole = readr::read_csv('data/womensrole.csv')
womensrole_glm <- glm(cbind(agree, disagree) ~ education + gender,
                        data = womensrole, 
                        family = binomial)
summary(womensrole_glm)
```

```{r basic_logistic_result, echo=FALSE, eval=FALSE}
tidy(womensrole_glm) %>% 
  kable_df()
```